Modal Abstractions for Virtualizing Memory Addresses

### 2024-07-10 examples.tex
+

$\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \textsf{entry\_val}_{\textsf{pfn}} \mapsto_{\textsf{id}} \textsf{\_} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val}_{\textsf{pfn}})  \urcorner  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$

@@

$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{entry\_val}_{\textsf{pfn}}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$
$\specline{ \forall_{i\in \textsf{0 ... 511} } \ldotp  \textsf{entry\_val }_{\textsf{pfn}} + \textsf{i * 8} \mapsto_{\textsf{id}} \textsf{\_} \ast  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val }_{\textsf{pfn}} + \textsf{i * 8})  }$

@@

$\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \textsf{entry\_val}_{\textsf{pfn}} \mapsto_{\textsf{id}} \textsf{\_} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \lnot(\textsf{entry\_present entry\_val}_{\textsf{pfn}})  \urcorner  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{entry\_val}_{\textsf{pfn}}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$
$\specline{ \forall_{i\in \textsf{0 ... 511} } \ldotp  \textsf{entry\_val }_{\textsf{pfn}} + \textsf{i * 8} \mapsto_{\textsf{id}} \textsf{\_} \ast  \textsf{qfrac} = 1 \leftrightarrow \lnot(\textsf{entry\_present entry\_val }_{\textsf{pfn}} + \textsf{i * 8})  }$

@@

$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \} \cup \{\textsf{entry\_val}_{\textsf{pfn}}) \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$

@@

$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{entry\_val}_{\textsf{pfn}}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{entry\_val}_{\textsf{pfn}}  }_{\rtv}$

@@

Looking at the update from Specfication Lines 3 to 15 in Figure \ref{walkpgdir}, we see that the removal of \textsf{entry} from the map $\Xi$ in the invariant to construct the \textsf{pte-ptsto} (Specfication Line 15 in Figure \ref{walkpgdir}) needed for executing the load instruction (Line 15 in Figure \ref{fig:calltopteinitialize}) to obtain the value of a page entry whose physical address (\textsf{entry}) kept in the register $\texttt{r8}$. This magical abstraction -- represented as $\Xi$ in the invariant definition -- provides us to construct this \textsf{pte-ptsto} out of an physical address is called \textit{identity-mappings}.
\subsubsection{Identity Mappings}
\label{subsec:identitymappings}
Most kernels maintain an invariant on their page tables that the virtual address of any page used for a page table lives at a virtual address
whose value is a constant offset from the physical address a practice sometimes referred to as \emph{identity mapping} (even though the physical-to-virtual translation
is typically not literally the identity function, but adding a constant offset).\footnote{Some kernels do this for all physical memory on the machine, simplifying interaction
with DMA devices.}
\begin{figure*}
\footnotesize
\todo[inline]{Ismail, look at this figure, something went weird (was already like this, I didn't add this)}
\begin{array}{l}
  \mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi,m)\stackrel{\triangle}{=}\\ \textsf{ASpace\_Lookup}_{\textsf{id}}(\ptablestore,\Xi,m) \ast 
  \bigast{(\vaddr, \textsf{paddr})\in \ptablestore}{\exists\;(\textsf{l4e, l3e, l2e, l1e, paddr})\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr\textsf{, l4e, l3e, l2e, l1e, paddr})} \\
  \bigast{(\paddr,\mathsf{T}) \in \Xi}{\exists\; (\textsf{qfrac, q, val,}\vaddr) \ldotp \ulcorner \vaddr = \paddr + \textsf{KERNBASE} \urcorner \ast  \underbrace{\fracghostmaptoken{\delta}{\vaddr}{\paddr}{\qfrac} }_\text{Ghost translation} \ast \underbrace{\paddr \mapsto_{\mathsf{p}}\{\textsf{qfrac}\}\; \vale}_\text{Physical location}}\\
   \forall_{\textsf{i}\in\textsf{0..511}} \ldotp \paddr + \textsf{i * 8} \mapsto_{\textsf{id,qfacc}} \textsf{mvalidc} \ast  \underbrace{ \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner}_\text{Entry validity}\ast \\
  \textsf{ where } \\
   \textsf{ASpace\_Lookup}_{\textsf{id}}(\ptablestore,\Xi,m) \stackrel{\triangle}{=} \lambda\textsf{ cr3val} \ldotp \; \exists \gammaPred \; \ldotp \ulcorner m \; !!\; \textsf{cr3val} = \textsf{Some } \gammaPred \urcorner \ast
   % \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \ast  \ownGhostpv\gammaPred{\authfull{\pvmapping\Xi}}
   \ptableabswalk{\delta,\ptablestore} \ast \pvmapping{\delta,\Xi}

@@

Concretely speaking, when we focus on the specfication of Line 15 in Figure \ref{fig:calltopteinitialize} in Figure \ref{fig:lookingupidpampping}, to read the value referenced by physical address \textsf{entry}, we need the guarantee
under the side condition (*)
\begin{equation*}
 \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner \tag{*}
\end{equation*}
assuring that looking the identity mapping for \textsf{entry} is safe under the subtle justification which equates the full ownership to the non/presence of the entry which can only be known when investigated in Line 21 in Figure \ref{fig:calltopteinitialize}.
\begin{comment}
 \begin{figure*}
   \footnotesize
   \[
 \begin{array}{l}
   \specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \textsf{\_} \ast \ulcorner \Xi !! \textsf{entry} = \textsf{Some V} \urcorner  }_{\rtv}  \\
\sqsubseteq \\
   \specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv} \\
\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv} \\
\mathsf{mov} \;   [ \mathsf{r8} ],\mathsf{rdi}\\
 \specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv} \\
\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv} \\
 \end{array}
 \]
 \vspace{-1em}
\caption{Peel-out the resources needed for the virtual-pte pointsto relation for physical address \textsf{entry} to be accessed in Line 7 in Figure \ref{fig:calltopteinitialize}. }
  \label{fig:lookingupidpampping}
 \end{figure*}
\end{comment}
 \subsubsection{Ensuring Presence of an Entry}
 After obtaining the identity mapping for \textsf{entry}, we are able to load the \textsf{entry\_val} into \textsf{rdi}, and check the presence bit through
 \[ \begin{array}{l}
   \mathsf{and }\; 0x1, \;rdi \\
   \mathsf{mov } \;\mathsf{rdi}, \; \mathsf{rax} \\
   \mathsf{cmp } \; 0x0, \;\mathsf{rax}
 \end{array}
 \] Lines (19-22) in Figure \ref{fig:calltopteinitialize}.
 Accessing the presence bit and checking the value allow  us to exploit the fact in the side condition (*) in such a way that we either take the allocation path in the non-presence case (Lines 19-44 in Figure \ref{fig:calltopteinitialize}) with the knowledge of full ownership ($\textsf{qfrac} = 1$) to set the entry with proper values for the bitfields, most imporantly frame referencing freshly allocated page and presence bit, or, take non-allocation path and jump to Line 45 with fractional permission for looking up the entry values.

 In case of absence, \textsf{pte\_initialize} is called (Line 33 in Figure \ref{fig:calltopteinitialize}) for a physical-zeroed page (utilizing the page-allocator's \textsf{kalloc} -- currently the only axiomatized call in the the proof of \textsf{pte\_initialize} (Line 9 in Figure \ref{pteinitializespec}, it is not part of \textsf{pte} (virtual-memory-management) module). 
\begin{figure}\footnotesize
  \begin{lstlisting}[mathescape]

$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
$\specline{\texttt{rax} \mapsto_{\textsf{r}} \textsf{\_} \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\_}\ast\ghostmaptoken{\delta{}s}{\rtv}{\delta} }_{\rtv}$
;;void pte_initialize(pte_t **table) {
...
;;set up the stack    
;;allocate a full zeroed page for 512 8-byte entries 
;;pte_t *local = kalloc();
callq  81 <pte_initialize+0x11>	7d: R_X86_64_PLT32	kalloc-0x4
mov    rax,-0x10[rbp] 
;;*table = local;
mov    -0x10[rbp],rax
mov    -0x8[rbp],rdi
mov    rax,[rdi]
$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\mathsf{nextpaddr}} \ast \textsf{nextpaddr} \mapsto_{\textsf{id}} \textsf{\_}\ast \forall_{i\in\textsf{0..511}} \textsf{nextpaddr + i * 8} \mapsto_{\textsf{id}} \textsf{\_} \ast\ghostmaptoken{\delta{}s}{\rtv}{\delta} }_{\rtv}$
;;clean up the stack
...
\end{lstlisting}
\vspace{-1em}
\caption{Allocating a physical page }
\label{pteinitializespec}
\end{figure}
Since we are using \textsf{pte\_initialize} for page-table address allocation, we must relate this newly allocated physical address to the identity mapping map $\Xi$ -- Specification Lines 34 and 38 in Figure \ref{fig:calltopteinitialize}. There are two important aspects on relating a freshly allocated page-table physical address to the identity-mapping which we focus in more detail separately in the specification of \textsf{pte\_initialize} in Figure \ref{fig:pteinitializespec}. First, one might have expected to see a physical pointsto for the fresly page-table address  ($\mathsf{nextpaddr} \mapsto_{\mathsf{p}} \mathsf{w64\_0}$) after the call to \textsf{pte\_initialize} but we do not have this physical pointsto resource. Second, we see the identity mapping for the entry frame ($\textsf{entry\_val}_{\textsf{pfn}}$), and all the knowledge on the existence of the identity mappings for both the frame and the page entries indexed from the frame (Line 38 in Figure \ref{fig:calltopteinitialize}) identity mappings exists, which are not part of the invariant yet, which could occur once the presence bit of the entry is set (Line 42 in Figure \ref{fig:calltopteinitialize}). Presence bit guard in the extended invariant shown in Figure \ref{fig:peraspaceinvariant_with_p2v_extension} \todo{we need to say something about the improtance of being present and bein inside the invariant.} allows us to preserve the \textit{soundness} of the table-traversal it eliminates all the entries (and tables) that are being under mutation for any other access other than mutation itself.
%%%%deprecated 
\begin{comment}
once the presence bit of the entry for which the invariant  the entry}first not appearing in the identity map (Specification Line 11 in Figure \ref{fig:pteinitializespec}), then as inserted into the identity-map with a value \textsf{Z} ($\Xi \cup \{\mathsf{nextpaddr}\}$) together with the fragmental capability  -- $\ulcorner \Xi \; !!\; \textsf{nextpaddr} = \textsf{ Some Z} \urcorner$ -- at Lines 13 and 14 in Figure \ref{fig:pteinitializespec} . The value type \textsf{Z}, unlike the value type \textsf{V} denoting the random value in Line 8 in Figure \ref{fig:calltopteinitialize}, denotes \textsf{zeroed} value.

    The first aspect is important for not invalidating the soundness: \textsf{pte\_initialize} extends the identity map with the freshly allocated \textsf{nextpaddr}, and must not reveal the ownership of the physical pointsto in the post-condition (Line 33 and 35 in Figure \ref{fig:calltopteinitialize}, Lines 22-23 in Figure \ref{fig:pteinitializespec}) because it would create unsound coexistence of the full-ownership of the physical pointsto resource (a hypothetical $\mathsf{nextpaddr} \mapsto_{\mathsf{p,1}} w0_64$) and fragmental (potentially full based on the side-condition *) physical pointsto resource that can be obtained from the identity map inside the extended invariant in Figure \ref{fig:peraspaceinvariant_with_p2v_extension} with the capability token $\ulcorner \Xi \;!!;\ \textsf{nextpaddr} = \textsf{ Some Z} \urcorner$. Although, the first aspect -- not explicitly revealing the physical pointsto resource for the freshly allocated page-table address but providing access through the identity map inside the invariant with access-capability tokens -- recovers the soundness argument expalined, we still need the knowledge that the fresly allocated address references to a \textit{zeroed} page. The invariant preserves this information through the constraint on relating the type (\textsf{T}) of mapped data -- \textsf{T = Z} for zeroed or \textsf{T = V} for random page value -- to the value of existentially quantified value -- \textsf{val} -- in Figure \ref{fig:peraspaceinvariant_with_p2v_extension}. So, for accessing the physical pointsto resource to the zeroed value referenced by \textsf{nextpaddr}, we need to have the fragmental ownership of the identity-map $\Xi \; !! \; \textsf{nextpaddr} = \textsf{ Some Z}$ which ensures and carries the \textit{being-zeroed} knowledge when we open and access the the existentially quantified value inside the invariant.
\end{comment}
%%%%%%%%%%
  \subsubsection{Kernels Utilizing \textsf{P2V}}
  \label{sec:p2v}
 Once we make sure that the frame of the page-table entry ($\textsf{entry}\mapsto\textsf{pfn}$)is backed with an allocated physical page, we must obtain the virtual address of the next level-page table (Line 48 in \ref{calltopteinitialize}). To do so, as for obtaining any virtual page-table address, we, again, need to open the invariant look up the \textit{identity mapping} for the physical address ($\textsf{entry}\mapsto \textsf{pfn}$).

This actually constitutes a very critical piece of the full page table walk verification, and we have verified the critical step for a small x86-64 kernel, which is the physical-to-virtual conversion (often appearing as a macro \texttt{P2V} inside the macro \textsf{ PTE\_PFN\_TO\_ADDR} in C source code). In our small kernel (Line 48 in Figure \ref{calltopteinitialize}), as in larger kernels, \texttt{P2V} is actually just addition, but the correctness of this simple instruction is quite subtle and relies on identity mappings explained in the previous section (Section \ref{subsec:identitymappings}) as a part of understanding our address-space invariant in Figure \ref{fig:peraspaceinvariant_with_p2v_extension}.

@@

$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{entry\_val}_{\textsf{pfn}}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{entry\_val}_{\textsf{pfn}}  }_{\rtv}$

>>>>

$\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \textsf{entry}_{\textsf{pfn}} \mapsto_{\textsf{id}} \textsf{\_} \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val}_{\textsf{pfn}})  \urcorner  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$

@@

$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{entry}_{\textsf{pfn}}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$
$\specline{ \forall_{i\in \textsf{0 ... 511} } \ldotp  \textsf{entry }_{\textsf{pfn}} + \textsf{i * 8} \mapsto_{\textsf{id}} \textsf{\_} \ast  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entry\_val }_{\textsf{pfn}} + \textsf{i * 8})  }$

@@

$\specline{ \textsf{entry} \mapsto_{\textsf{id}} \textsf{\_} \ast \textsf{entry}_{\textsf{pfn}} \mapsto_{\textsf{id}} \textsf{\_} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \lnot(\textsf{entry\_present entry\_val}_{\textsf{pfn}})  \urcorner  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \mathsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{entry}_{\textsf{pfn}}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{nextpaddr}  }_{\rtv}$
$\specline{ \forall_{i\in \textsf{0 ... 511} } \ldotp  \textsf{entry }_{\textsf{pfn}} + \textsf{i * 8} \mapsto_{\textsf{id}} \textsf{\_} \ast  \textsf{qfrac} = 1 \leftrightarrow \lnot(\textsf{entry\_present entry\_val }_{\textsf{pfn}} + \textsf{i * 8})  }$

@@

$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \} \cup \{\textsf{entry}_{\textsf{pfn}}) \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$

@@

$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \mathsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \lnot(\textsf{entry\_present entry\_val})  \urcorner }_{\rtv}$
$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{entry}_{\textsf{pfn}}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{entry}_{\textsf{pfn}}  }_{\rtv}$

@@

The most crucial aspect of page-table entry access relies on the subtle interplay between updating the shared page tables, and the not violating the memory mappings. Concretely speaking, accessing to the entry -- whose address is kept in the register $\texttt{r8}$ -- in Line 15 in Figure \ref{fig:calltopteinitialize} requires a memory mapping that is logically provided through the update from Specfication Lines 3 to 15 in Figure \ref{walkpgdir} on the client side by the removal of \textsf{entry} from the map $\Xi$ in the invariant to construct the \textsf{vpte-ptsto} (Specfication Line 15 in Figure \ref{walkpgdir}). This magical abstraction -- represented as $\Xi$ in the invariant definition -- provides us to construct this \textsf{pte-ptsto} out of an physical address is called \textit{identity-mappings}.
\subsubsection{Identity Mappings}
\label{subsec:identitymappings}
Most kernels maintain an invariant on their page tables that the virtual address of any page used for a page table lives at a virtual address
whose value is a constant offset from the physical address a practice sometimes referred to as \emph{identity mapping} (even though the physical-to-virtual translation
is typically not literally the identity function, but adding a constant offset).\footnote{Some kernels do this for all physical memory on the machine, simplifying interaction
with DMA devices.}
\begin{figure*}
\footnotesize
\[
\begin{array}{l}
  \mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi,m)\stackrel{\triangle}{=}\\ \textsf{ASpace\_Lookup}_{\textsf{id}}(\ptablestore,\Xi,m) \ast 
  \bigast{(\vaddr, \textsf{paddr})\in \ptablestore}{\exists\;(\textsf{l4e, l3e, l2e, l1e, paddr})\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr\textsf{, l4e, l3e, l2e, l1e, paddr})} \\
  \bigast{(\paddr,\mathsf{T}) \in \Xi}{\exists\; (\textsf{qfrac, q, val,}\vaddr) \ldotp \ulcorner \vaddr = \paddr + \textsf{KERNBASE} \urcorner \ast  \underbrace{\fracghostmaptoken{\delta}{\vaddr}{\paddr}{\qfrac} }_\text{Ghost translation} \ast \underbrace{\paddr \mapsto_{\mathsf{p}}\{\textsf{qfrac}\}\; \vale}_\text{Physical location}}\\
   \forall_{\textsf{i}\in\textsf{0..511}} \ldotp \paddr + \textsf{i * 8} \mapsto_{\textsf{id,qfacc}} \textsf{mvalidc} \ast  \underbrace{ \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner}_\text{Entry validity}\ast \\
  \textsf{ where } \\
   \textsf{ASpace\_Lookup}_{\textsf{id}}(\ptablestore,\Xi,m) \stackrel{\triangle}{=} \lambda\textsf{ cr3val} \ldotp \; \exists \gammaPred \; \ldotp \ulcorner m \; !!\; \textsf{cr3val} = \textsf{Some } \gammaPred \urcorner \ast
   % \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \ast  \ownGhostpv\gammaPred{\authfull{\pvmapping\Xi}}
   \ptableabswalk{\delta,\ptablestore} \ast \pvmapping{\delta,\Xi}

@@

Concretely speaking, going back to Line 15 in Figure \ref{fig:calltopteinitialize}, to read the value referenced by physical address \textsf{entry} while preserving the soundness of memory mappings, our extended invariant introduces the side condition (*)
\begin{equation*}
 \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner \tag{*}
\end{equation*}
assuring that looking the identity mapping for \textsf{entry} is safe under the subtle justification which equates the full ownership to the non/presence of the entry which can only be known when investigated in Line 21 in Figure \ref{fig:calltopteinitialize}.
\begin{comment}
 \begin{figure*}
   \footnotesize
   \[
 \begin{array}{l}
   \specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \textsf{\_} \ast \ulcorner \Xi !! \textsf{entry} = \textsf{Some V} \urcorner  }_{\rtv}  \\
\sqsubseteq \\
   \specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv} \\
\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv} \\
\mathsf{mov} \;   [ \mathsf{r8} ],\mathsf{rdi}\\
 \specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv} \\
\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv} \\
 \end{array}
 \]
 \vspace{-1em}
\caption{Peel-out the resources needed for the virtual-pte pointsto relation for physical address \textsf{entry} to be accessed in Line 7 in Figure \ref{fig:calltopteinitialize}. }
  \label{fig:lookingupidpampping}
 \end{figure*}
\end{comment}
 \subsubsection{Ensuring Presence of an Entry}
 After obtaining the identity mapping for \textsf{entry}, we are able to load the \textsf{entry\_val} into \textsf{rdi}, and check the presence bit through
 \[ \begin{array}{l}
   \mathsf{and }\; 0x1, \;rdi \\
   \mathsf{mov } \;\mathsf{rdi}, \; \mathsf{rax} \\
   \mathsf{cmp } \; 0x0, \;\mathsf{rax}
 \end{array}
 \] Lines (19-22) in Figure \ref{fig:calltopteinitialize}.
 Accessing the presence bit and checking the value allow  us to exploit the fact in the side condition (*) in such a way that we either take the allocation path in the non-presence case (Lines 19-44 in Figure \ref{fig:calltopteinitialize}) with the knowledge of full ownership ($\textsf{qfrac} = 1$) to set the entry with proper values for the bitfields, most imporantly frame referencing freshly allocated page and presence bit, or, take non-allocation path and jump to Line 45 with fractional permission for looking up the entry values.

 In case of absence, \textsf{pte\_initialize} is called (Line 33 in Figure \ref{fig:calltopteinitialize}) for a physical page (utilizing the page-allocator's \textsf{kalloc} -- currently the only axiomatized call in the the proof of \textsf{pte\_initialize} (Line 9 in Figure \ref{pteinitializespec}, it is not part of \textsf{pte} (virtual-memory-management) module). 
\begin{figure}\footnotesize
  \begin{lstlisting}[mathescape]

$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
$\specline{\texttt{rax} \mapsto_{\textsf{r}} \textsf{\_} \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\_}\ast\ghostmaptoken{\delta{}s}{\rtv}{\delta} }_{\rtv}$
;;void pte_initialize(pte_t **table) {
...
;;set up the stack    
;;allocate a full zeroed page for 512 8-byte entries 
;;pte_t *local = kalloc();
callq  81 <pte_initialize+0x11>	7d: R_X86_64_PLT32	kalloc-0x4
mov    rax,-0x10[rbp] 
;;*table = local;
mov    -0x10[rbp],rax
mov    -0x8[rbp],rdi
mov    rax,[rdi]
$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\mathsf{nextpaddr}} \ast \textsf{nextpaddr} \mapsto_{\textsf{id}} \textsf{\_}\ast \forall_{i\in\textsf{0..511}} \textsf{nextpaddr + i * 8} \mapsto_{\textsf{id}} \textsf{\_} \ast\ghostmaptoken{\delta{}s}{\rtv}{\delta} }_{\rtv}$
;;clean up the stack
...
\end{lstlisting}
\vspace{-1em}
\caption{Allocating a physical page }
\label{pteinitializespec}
\end{figure}
Since we are using \textsf{pte\_initialize} for page-table address allocation, we must relate this newly allocated physical address to the identity mapping map $\Xi$ -- Specification Lines 34 and 38 in Figure \ref{fig:calltopteinitialize}. Focusing on the specification of \textsf{pte\_initialize} separately in Figure \ref{fig:pteinitializespec}, we right immediately realize that instead of seeing see a physical pointsto for the fresly page-table address (e.g. $\mathsf{nextpaddr} \mapsto_{\mathsf{p}} \mathsf{w64\_0}$) deliberately in the post-conditoin in Lines 15-16, we observe a full-ownership token representing the knowledge that a frame and all the entries indexed from this frame are freshly allocated with full-ownership to be a part of the identity map, $\Xi$. The soundness argument of this specification relies on the fact that these freshly allocated resources is part of an entry construction that has not been completed yet: the presence bit is set (Line 42 in Figure \ref{fig:calltopteinitialize}) after these freshly allocated resources are incorporated to the entry construction (Line 41 in Figure \ref{fig:calltopteinitialize}). In other words, the side condition, (*), acknowledges any access to the entry with these resources as \textit{invalid}, as the entry is revealed to shared-accesses when the presence bit is set. 
  \subsubsection{Kernels Utilizing \textsf{P2V}}
  \label{sec:p2v}
 Once we reach to the certain knowledge of having an entry with a frame referencing to an allocated resource resides inside the identity mappings which can already be known (not taking the allocation path in (Lines 21-53 in Figure \ref{fig:calltopteinitialize}) or by ensuring the allocated frame is made present to shared accesses, then put into identity mappings -- logical update from Specification Lines 43-47 to 49-52 putting $\textsf{entry}_{\textsf{pfn}}$ into $\Xi$ in Figure \ref{fig:calltopteinitialize}), no matther how it is constructed, we can utilize this knowledge to convert this frame address into an virtual address of the next page table through, again, identity mappings -- Line 57 in Figure \ref{fig:calltopteinitialize} specified in Figure \ref{fig:p2v}.

This actually constitutes a very critical piece of the full page table walk verification, and we have verified the critical step for a small x86-64 kernel, which is the physical-to-virtual conversion (often appearing as a macro \texttt{P2V} inside the macro \textsf{ PTE\_PFN\_TO\_ADDR} in C source code). In our small kernel (Line 57 in Figure \ref{calltopteinitialize}), as in larger kernels, \texttt{P2V} is actually just addition, but the correctness of this simple instruction is quite subtle and relies on identity mappings explained in the previous section (Section \ref{subsec:identitymappings}) as a part of understanding our address-space invariant in Figure \ref{fig:peraspaceinvariant_with_p2v_extension}.

@@

$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val}  \urcorner }_{\rtv}$
$\specline{ \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{entry}_{\textsf{pfn}}  \ast \texttt{rax} \mapsto_{\textsf{r}} \textsf{entry}_{\textsf{pfn}} \ast \textsf{entry}_{\textsf{pfn}} \mapsto_{\textsf{id}} \textsf{\_}  }_{\rtv}$
